Existing Model

### 2022-11-07 DataSetup.Rmd
Updated code for difference score model, updated plottimng code.

Load Other Variables
```{r}
#Puberty
#Raw scores, youth
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(5:38) %>% 
  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
  mutate_at(c(5:28), as.numeric)

@@

#Cognition Scores (Age-Corrected)
cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(src_subject_id,interview_age, eventname, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected, nihtbx_totalcomp_rawscore, nihtbx_totalcomp_cs, nihtbx_totalcomp_itmcnt) %>% 
  mutate_at(c(2, 4:9), as.numeric)

@@

social <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(-c(subjectkey, interview_date, interview_age, gish2_device)) %>% 
  mutate_at(c(4:16), as.numeric)

@@

# Multiwave
youth_pds_multiwave <- youth_pds %>% 
  filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
  mutate(youth_sum = case_when(sex == "M" ~ #male
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_m4_y + pds_m5_y),  
                         sex == "F" ~ #female
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_f4_2_y + pds_f5_y))) %>% 
  mutate(youth_mean = youth_sum/5)  %>% 
  select(src_subject_id, sex, youth_sum, youth_mean)

youth_pds_baseline <- youth_pds_multiwave %>% 
  filter(eventname == "baseline_year_1_arm_1")

youth_pds_followup <- youth_pds_multiwave %>% 
  filter(eventname == "2_year_follow_up_y_arm_1")

@@

#Baseline

parent_pds_multiwave <- parent_pds %>% 
    filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
    mutate(parent_sum = case_when(sex == "M" ~ #male
                           (pds_1_p + pds_2_p + pds_3_p + pds_m4_p + pds_m5_p),  
                         sex == "F" ~ #female
                           (pds_1_p + pds_2_p + pds_3_p + pds_f4_p + pds_f5b_p))) %>% 
  mutate(parent_mean = parent_sum/5) %>% 
  select(src_subject_id, sex, parent_sum, parent_mean)

parent_pds_baseline <- parent_pds_multiwave %>% 
    filter(eventname == "baseline_year_1_arm_1")

parent_pds_followup <- parent_pds_multiwave%>% 
    filter(eventname == "2_year_follow_up_y_arm_1")

@@

#Baseline
cognition_baseline <- cognition %>% 
  filter(eventname == "baseline_year_1_arm_1")

#Follow-Up
cognition_followup <- cognition %>% 
  filter(eventname == "2_year_follow_up_y_arm_1")

#Multiwave
cognition_multiwave <- cognition %>% 
  filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1")
```

@@

corrected_followup_brainage$src_subject_id <- smri_followup$src_subject_id

@@

#Save samples
save(brainage_pds_followup, file="BrainagePDSFollowUp.Rda")
save(sample_1_followup, file="Sample1FollowUp.Rda")
save(sample_2_followup, file = "Sample2FollowUp.Rda")
```

>>>>

Load Other Variables
```{r}
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/BaselineBrainAgeCorrected.Rda")
load("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/FollowUpBrainAgeCorrected.Rda")

#Puberty
#Raw scores, youth
youth_pds <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_ypdms01.txt") %>%
  filter(!collection_title=="collection_title") %>% 
  select(5:38) %>% 
  select(-c(interview_date, pds_remote___1, pds_remote___2, pds_remote___3, pds_remote___4, pds_device)) %>% 
  mutate_at(c(5:28), as.numeric)

@@

#Cognition Scores (Age-Corrected)
cognition <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_tbss01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  select(src_subject_id,interview_age, eventname, nihtbx_fluidcomp_agecorrected, nihtbx_cryst_agecorrected, nihtbx_totalcomp_agecorrected, nihtbx_totalcomp_uncorrected) %>% 
  mutate_at(c(2, 4:7), as.numeric)

@@

gish <- rio::import("/Volumes/devbrainlab/ABCD_Data/ABCD4pt0/abcd_gish2y01.txt") %>%
  filter(!collection_title=="collection_title")  %>% 
  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
   select(src_subject_id, sex, contains("gish2")) %>% 
  select(-c(gish2_device, abcd_gish2y01_id))%>% 
  mutate_at(c(3:15), as.numeric)

@@

# Multiwave
youth_pds_multiwave <- youth_pds %>% 
  filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
  mutate(youth_sum = case_when(sex == "M" ~ #male
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_m4_y + pds_m5_y),  
                         sex == "F" ~ #female
                           (pds_ht2_y + pds_bdyhair_y + pds_skin2_y + pds_f4_2_y + pds_f5_y))) %>% 
  mutate(youth_mean = youth_sum/5)  %>% 
  select(src_subject_id, eventname, sex, youth_sum, youth_mean)

youth_pds_baseline <- youth_pds_multiwave %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  select(-eventname)

youth_pds_followup <- youth_pds_multiwave %>% 
  filter(eventname == "2_year_follow_up_y_arm_1") %>% 
  select(-eventname)

@@

#Baseline
parent_pds_multiwave <- parent_pds %>% 
    filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1") %>% 
    mutate(parent_sum = case_when(sex == "M" ~ #male
                           (pds_1_p + pds_2_p + pds_3_p + pds_m4_p + pds_m5_p),  
                         sex == "F" ~ #female
                           (pds_1_p + pds_2_p + pds_3_p + pds_f4_p + pds_f5b_p))) %>% 
  mutate(parent_mean = parent_sum/5) %>% 
  select(src_subject_id, eventname, sex, parent_sum, parent_mean)

parent_pds_baseline <- parent_pds_multiwave %>% 
    filter(eventname == "baseline_year_1_arm_1") %>% 
  select(-eventname)

parent_pds_followup <- parent_pds_multiwave%>% 
    filter(eventname == "2_year_follow_up_y_arm_1")%>% 
  select(-eventname)

@@

#Baseline
cognition_baseline <- cognition %>% 
  filter(eventname == "baseline_year_1_arm_1") %>% 
  select(-eventname) %>%
  mutate(cog_zscore = scale(nihtbx_totalcomp_uncorrected))

#Follow-Up
cognition_followup <- cognition %>% 
  filter(eventname == "2_year_follow_up_y_arm_1")%>% 
  select(-eventname)

#Multiwave
cognition_multiwave <- cognition %>% 
  filter(eventname == "baseline_year_1_arm_1" | eventname == "2_year_follow_up_y_arm_1")

#Social/GISH
gish[gish== 777] <- NA
gish[gish == 999] <- NA

# sum scores
gish %>% 
  mutate(gishSum = rowSums(across(c(3:15)), na.rm=T))

@@

corrected_followup_brainage$src_subject_id <- smri_followup$src_subject_id
```

@@

#Save samples
save(brainage_pds_followup, file="BrainagePDSFollowUp.Rda")
save(sample_1_followup, file="Sample1FollowUp.Rda")
save(sample_2_followup, file = "Sample2FollowUp.Rda")

#Add Gish to follow-up
```

Attach maturity variables to ABCD model estimates
```{r}
abcd_brainage_baseline <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp/ABCDBaselineBrainAgeCorrected.Rda")
abcd_brainage_baseline$src_subject_id <- sample_2_baseline$src_subject_id

abcd_brainage_baseline <- abcd_brainage_baseline[, c(6,1:5)]

abcd_baseline <- sample_2_baseline %>% 
  select(1, 7:17)

abcd_baseline <- left_join(abcd_baseline, abcd_brainage_baseline)

save(abcd_baseline, file="ABCDBaselineSample.Rda")

#double check that sample 2s remain consistent

abcd_baseline_ids <- abcd_brainage_baseline$src_subject_id

og_sample_2 <- rio::import("/Volumes/devbrainlab/Lucy/BrainAGE/fyp_old/Sample2Baseline.Rda")
og_sample_2_ids <- og_sample_2$src_subject_id

id_compare<-data.frame(abcd_baseline_ids,og_sample_2_ids)
sum(ifelse(id_compare$abcd_baseline_ids==id_compare$og_sample_2_ids,1,0))

#they're the same!

### 2022-11-07 DifferenceScores.Rmd
Updated code for difference score model, updated plottimng code.

smri_difference_calculated_scaled<- smri_difference_calculated_subset[,-(1:2)] %>% sapply(`/`, smri_difference_calculated_subset[,2])

smri_difference_calculated_scaled$src_subject_id <- smri_difference_calculated_subset$src_subject_id

smri_difference_calculated_scaled$interview_age_diff <- smri_difference_calculated_subset$interview_age.y_diff

smri_difference_calculated_scaled <- smri_difference_calculated_scaled
select(src_subject_id, interview_age_diff, everything())

# Should we be predicting the change in age?
set.seed(42)
smri_difference_calculated <- smri_difference_calculated %>% 
#split for testing
select(-c(src_subject_id))

>>>>

smri_difference_calculated_subset[, -(1:2)] <- sweep(smri_difference_calculated_subset[, -(1:2)], 1, smri_difference_calculated_subset[, 2], "/")

# testing
#difference_df_scaled<- difference_df[,-(1:2)] %>% sapply(`/`, difference_df[,2])

#smri_difference_calculated_scaled<- smri_difference_calculated_subset[,-(1:2)] %>% sapply(`/`, smri_difference_calculated_subset[,2]) 
#try lapply instead

#difference_df_scaled<- difference_df %>%  
#  mutate(across(everything(), -c(difference1, difference2) ~ . / !! difference2))

#difference_df_scaled<- difference_df %>% 
 # mutate(across(!difference1 & !difference2), ~ . / !! difference2)

#difference_df_scaled<- difference_df %>% 
 # mutate(across(.cols = -c(difference1, difference2), . / difference2(., na.rm = FALSE)))

#difference_df_scaled<- difference_df %>%
#  mutate_at(vars(-difference1 -difference2),
  #          ~ . / !! difference2
#  )

#difference_df_scale <- difference_df
#difference_df_scale[, -(1:2)] <- sweep(difference_df_scale[, -(1:2)], 1, difference_df_scale[, 2], "/")

####

#smri_difference_calculated_scaled$src_subject_id <- smri_difference_calculated_subset$src_subject_id

#smri_difference_calculated_scaled$interview_age_diff <- smri_difference_calculated_subset$interview_age.y_diff

# Should we be predicting the change in age?

#probably filter out people w/o age_diff scores
smri_difference_calculated_filtered <- smri_difference_calculated_subset %>% 
  filter(!is.na(interview_age.y_diff))

#calculate midpoint age, use change scores to predict midpoint
smri_difference_calculated_filtered <- smri_difference_calculated_filtered %>% 
  left_join(smri_difference %>%  select(src_subject_id, interview_age.x, interview_age.y)) %>% 
  rowwise() %>% 
  mutate(age_midpoint = mean(c(interview_age.x, interview_age.y))) %>% 
  ungroup()

smri_difference_calculated_filtered <- smri_difference_calculated_filtered %>% 
  ungroup()

save(smri_difference_calculated_filtered, file="smri_difference_scores.Rda")

# select just subject_id, brain columns, age midpoint

#load("smri_difference_scores.Rda")

#1000 people for training

#Split sample
smp_size_diff <- floor(0.13 * nrow(smri_difference_calculated_filtered)) #training sample of 1000 participants

## set the seed to make your partition reproducible
set.seed(123)
sample_ind_diff <- sample(seq_len(nrow(smri_difference_calculated_filtered)), size = smp_size_diff)

sample_train_diff <- smri_difference_calculated_filtered[sample_ind_diff, ]
sample_test_diff <- smri_difference_calculated_filtered[-sample_ind_diff, ]

# select just subject_id, brain columns, age midpoint
sample_train_diff <- sample_train_diff %>% 
select(-c(src_subject_id, interview_age.y_diff, interview_age.x, interview_age.y))

save(sample_train_diff, file="sample_train_diff.Rda")
save(sample_test_diff, file="sample_test_diff.Rda")

### 2022-11-07 ExisitingModel.Rmd
Updated file path to reflect submodule
Updated code for difference score model, updated plottimng code.

#Install XGboost & tidymodels
packageUrl <- "https://cran.r-project.org/src/contrib/Archive/xgboost/xgboost_1.0.0.2.tar.gz"
# You then install this version of the package using
install.packages(packageUrl, repos = NULL, type = 'source')

install.packages("tidymodels")

@@

xgb_mod <- readRDS(
  file = here::here(
    "xgboost_9to19_brain_age_mod.rds"))
```

Predict Brain Age
```{r}
brain_age_df <-
  xgb_mod %>%
  predict(new_data = renamed_smri) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = renamed_smri$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

@@

Evaluate accuracy
```{r}
# compute common performance metrics: mae, rsq, rmse
brain_age_df %>%
  metrics(truth = truth, estimate = .pred)

@@

# create bias correct data frame
brain_age_corrected_df <- brain_age_df %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - bias_intercept) / bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

@@

#mine
brain_age_corrected_df %>%
  metrics(truth = truth, estimate = corrected_pred)

@@

save(brain_age_df, file="BaselineBrainAge.Rda")
save(brain_age_corrected_df, file="BaselineBrainAgeCorrected.Rda")

>>>>

#Install XGboost & tidymodels
#packageUrl <- "https://cran.r-project.org/src/contrib/Archive/xgboost/xgboost_1.0.0.2.tar.gz"
# You then install this version of the package using
#install.packages(packageUrl, repos = NULL, type = 'source')

#install.packages("tidymodels")

@@

xgb_mod <- readRDS(
  file = here::here(
    "DevelopmentalBrainAge/model/xgboost_9to19_brain_age_mod.rds"))
```

Predict Brain Age
```{r}
brain_age_df_baseline <-
  xgb_mod %>%
  predict(new_data = renamed_smri_baseline) %>%
  mutate(
    # provide the chronological age at time of scan
    truth = renamed_smri_baseline$interview_age
  ) %>%
  # compute the brain age gap by subtracting chronological age from prediction
  mutate(gap = .pred - truth)

@@

Evaluate accuracy
```{r}
# compute common performance metrics: mae, rsq, rmse
brain_age_df_baseline %>%
  metrics(truth = truth, estimate = .pred)

@@

# create bias correct data frame
brain_age_corrected_df_baseline <- brain_age_df_baseline %>%
  mutate(
    # corrected brain age prediction
    corrected_pred =  (.pred - bias_intercept) / bias_slope
  ) %>%
  # corrected corrected brain age gap
  mutate(corrected_gap = corrected_pred - truth)

@@

#mine
brain_age_corrected_df_baseline %>%
  metrics(truth = truth, estimate = corrected_pred)

@@

save(brain_age_df_baseline, file="BaselineBrainAge.Rda")
save(brain_age_corrected_df_baseline, file="BaselineBrainAgeCorrected.Rda")
